

\chapter{The Machines}

\dasm supports specific machines and processors through the provision of companion source-code files that can assist with programming each platform. These files are located in the \mono{machines} subdirectory. Support is provided for...

\begin{itemize}
\item \nameref{machine:atari2600}
\item Atari 7800
\item Channel F
\item 68hc11
\item 68hc908
\end{itemize}


\section{Atari 2600}
\label{machine:atari2600}

The Atari 2600 is a game console from 1978 that uses a 6507 processor. This processor is similar to the 6502 processor (supported by \dasm). The difference in the processors is the number of hardware address lines on the chips; these being 16 on the 6502, and 12 on the 6507. Thus, the 6502 can directly address 64K of memory and the 6507 only 4K of memory.  From the point of view of \dasm, the machines are identical, as the 6502 and 6507 share a common instruction set.

For programming the Atari 2600, use \mono{PROCESSOR 6502} at the start of your program.

\subsection{Support Files}

The Atari 2600 is explicitly supported with two files generally included in most programs for that machine.

\subsubsection{\mono{vcs.h}}

Contains the standardised register definitions for the \mono{RIOT} and \mono{TIA} chips, defined with uninitialised segments. The implementation allows relocation of the \mono{TIA} base address to a shadow register address.

\subsubsection{\mono{macro.h}}
\label{support6502}

Contains some useful macros.

\section{HD6303}

\subsection{Broken Opcodes Bug}
\label{changelog:20200903bug}
The \mono{AIM}, \mono{OIM}, \mono{EIM}, and \mono{TIM} opcodes are broken. These instructions are \textbf{three} bytes long according to the data sheets, yet in \dasm they are
treated as \textbf{two} byte instructions.

These instructions are supposed
to work as follows:

\begin{code}
	tim	#$10,$C2
	tim	#$80,$00,x
\end{code}

So there's an immediate value \textbf{and} a zero-page address
for these instructions! \dasm, however, only accepts these:

\begin{code}
	tim	$10
	tim	$10,x
\end{code}

These opcodes simply don't fit into the "regular pattern" of 8-bit CPUs
we deal with in \dasm. Fixing this will require changes to
\dasm beyond just fixing the instruction table, and the parser code is not even remotely ready for this.

The "workaround" for now is to use macros instead of the
actual instructions, see \mono{../test/broken6303hack.asm} for
details.


\section{Channel F}
\label{changelog:20200906f8}

\subsection{Processor selection}

With DASM, the target CPU is selected with the PROCESSOR
directive inside the source file that should be assembled.
The F8 CPU is selected like this:

\begin{usage}
	processor f8
	processor F8	; case insensitive
\end{usage}

\subsection{Expressions with parentheses}


Some of DASM's backends, for instance the one for the 6502,
don't allow parentheses in expressions that are part of a
mnemonic's operand, because parentheses are used in the 
6502's assembly language to denote indirect addressing.
Instead, you have to use brackets.

This is not the case with the F8 backend. Both parentheses
and brackets can be used everywhere, so the following lines
are parsed and assembled correctly:

\begin{code}
        as      (2+2)*2 ; Assembles to $c8
        as      [2+2]*2 ; Assembles to $c8
\end{code}

\subsection{Data definition directives}


Since \mono{DS} is an F8 instruction (decrement scratchpad register),
the \mono{DS} directive isn't available anymore if \dasm assembles
F8 code. Instead, use the \mono{RES} directive, which works just like
the \mono{DS} directive:

\begin{code}
        ds.b    4,$33   ; Would assemble to $33 $33 $33 $33,
                        ; but isn't available in F8 mode
        res.b   4,$33   ; Assembles to $33 $33 $33 $33
\end{code}
        
Of course \mono{RES.W} and \mono{RES.L} do exist as well.


For source code compatibility with \mono{f8tool} (another F8 assembler), some
additional data definition directives are available : \mono{DB}, \mono{DW} and \mono{DD}.
These work just like \mono{DC.B}, \mono{DC.W} and \mono{DC.L}:

\begin{code}
        dc.b    $f8     ; Assembles to $f8
        db      $f8     ; Assembles to $f8
\end{code}

\subsection{Special register names}

For some of the special registers, multiple names are accepted:

\begin{table}[H]
\begin{tabularx}{\linewidth}{cl}
\toprule
\textbf{Register}&\textbf{Accepted Names}\\
\hline

\mono{DC0}&\mono{DC},\mono{DC0}\\
\mono{PC0}&\mono{P0},\mono{PC0}\\
\mono{PC1}&\mono{P}, \mono{PC1}\\
\mono{J}&\mono{J}, Any expression that evaluates to 9\\
&(This may seem strange, but J is really\\
&just an alias for scratchpad register 9)\\
\hline
\end{tabularx}
\end{table}                        
                        
The names \mono{DC}, \mono{P0}, \mono{P} and \mono{J} are standard syntax, the other forms
have been introduced for compatibility with other assemblers.

Thus, the following lines assemble all correctly:

\begin{code}
        lr      h,dc    ; Assembles to $11
        lr      h,dc0   ; Assembles to $11
        lr      p0,q    ; Assembles to $0d
        lr      pc0,q   ; Assembles to $0d
        lr      p,q     ; Assembles to $09
        lr      pc1,q   ; Assembles to $09
        lr      w,j     ; Assembles to $1d
        lr      w,3*3   ; Assembles to $1d
\end{code}
        
\subsection{Scratchpad register access}

There are several ways to access scratchpad registers:

\begin{table}[H]
\begin{tabularx}{\linewidth}{ll}
\toprule
\textbf{Access Mode}&\textbf{Accepted Syntax}\\
\hline
\\
Direct access to registers \mono{0..11}&Any expression that evaluates to \mono{0..11}\\
Access via \mono{ISAR}&\mono{S}, \mono{(IS)}, any expression that evaluates to \mono{12}\\
Access via \mono{ISAR}, \mono{ISAR} incremented&\mono{I}, \mono{(IS)+}, any expression that evaluates to \mono{13}\\
Access via \mono{ISAR}, \mono{ISAR} decremented&\mono{D}, \mono{(IS)-}, any expression that evaluates to \mono{14}\\
\hline
\end{tabularx}
\end{table}

The \mono{(IS)}, \mono{(IS)+} and \mono{(IS)-} forms are not standard syntax and have
been mainly introduced for compatibility with \mono{f8tool}.


For some of the directly accessible scratchpad registers aliases exist:

\begin{table}[H]
\begin{tabularx}{\linewidth}{cc}
\toprule
\textbf{Register}&\textbf{Alias Name}\\
\hline
\mono{9}&\mono{J}\\
\mono{10}&\mono{HU}\\
\mono{11}&\mono{HL}\\
\hline
\end{tabularx}
\end{table}
        
Originally, \mono{J} was only used with the \mono{LR} instruction when accessing the
flags, but since \mono{J} is just an alias for register \mono{9}, \mono{J} can also be used
in normal scratchpad register operations.

The following lines assemble all correctly

\begin{code}
        xs      2+2             ; Assembles to $e4
        xs      s               ; Assembles to $ec
        xs      (is)            ; Assembles to $ec
        xs      12              ; Assembles to $ec
        xs      i               ; Assembles to $ed
        xs      (is)+           ; Assembles to $ed
        xs      13              ; Assembles to $ed
        xs      d               ; Assembles to $ee
        xs      (is)-           ; Assembles to $ee
        xs      14              ; Assembles to $ee
        xs      9               ; Assembles to $e9
        xs      j               ; Assembles to $e9
        xs      hu              ; Assembles to $ea
        xs      hl              ; Assembles to $eb
\end{code}


\subsection{No instruction optimizations are done}


The assembler doesn't optimize instructions where a smaller
instruction could be used. It won't optimize between
\mono{OUT/OUTS}, \mono{IN/INS} and \mono{LI/LIS}.

For instance, the following line assembles to \mono{\$20 \$00}, even
though the \mono{LIS} instruction could be used, which would need
only one byte.

\begin{code}
        li      0       ; Assembles to $20 $00
\end{code}
