\index{Change Log}
\section*{Change Log}
\label{section:changelog}

This section lists recent changes to the document, with most recent entries first. Each item line hyperlink to the relevant place in the document.

\subsection*{Changes}

\begin{itemize}

\item[]\mono{2020.08.24}
\item \hyperref[changelog:20200824const]{Removed 0x for hexadecimal. Apparently recently removed from the code?}
\item \hyperref[changelog:20200824license]{Upgraded license to GPL v3 to be compatible with .STY files used for generation of this manual}
\item \hyperref[changelog:20200824processor]{Fixed table of processor types to explicitly list PROCESSOR values}
\item \hyperref[changelog:20200824rangebug]{Documented dasm bug in constants signed range checking for 8-bit operands}
\item \hyperref[changelog:20200824const]{Constants and Numbers duplicate sections merged.}
\item \hyperref[changelog:20200824arithmetic]{Corrected error in table - || is logical-OR and \&\& is logical-AND when used in expressions. They will return 0 or 1 results.}
\item \hyperref[changelog:20200824org]{Clarified the ORG/RORG usage in the \nameref{flag:outputformat} table, in regards to the requirement that data must be in order in initialised segments}
\item \hyperref[changelog:20200824error]{Revamped the comment about error types}
\item \hyperref[changelog:20200824passes]{Mentioned default number of passes}
\item \hyperref[changelog:20200824developers]{Clarified the -d option}
\item \hyperref[changelog:20200824sourcefile]{Clarified commmand-line format}
\item \hyperref[changelog:specialcase]{Added note about special-casing of negative operands}
\item \hyperref[changelog:alternateunary]{Added alternates to unary operator table}
\item \hyperref[section:numberformat]{Added section describing \nameref{section:numberformat}}

\item[]\mono{2020.08.23}
\item \hyperref[changelog:20200823colon]{Added some usage notes about colon in label names and why it can be advantageous. I still hate it.}
\item \hyperref[section:changelog]{Added this change log section.}
\end{itemize}


